False Positive Probabilities for all Kepler Objects of Interest: \\ \nvalnew\ newly validated planets and \nfpnew\ likely false positives

### 2015-11-23 ms.tex
intro work; references

\newcommand{\nbadphot}{47} %EmptyPhotometry or MissingKOI
\newcommand{\nbadphotFP}{34}
\newcommand{\nbadphotCAND}{13}
\newcommand{\nbadrowemcmc}{216}
\newcommand{\nbadstellar}{98} %MissingStellar
\newcommand{\nbadsec}{162} %NoWeakSecondary
\newcommand{\nattempted}{8141}
\newcommand{\ntryfail}{112}
\newcommand{\nbadtrapfit}{32}
\newcommand{\nbadroche}{52}
\newcommand{\nbadother}{28}
\newcommand{\nbadotherFP}{22}
\newcommand{\nbadotherCAND}{5}
\newcommand{\nbadotherCONF}{1}

@@

\newcommand{\kepler}{\textit{Kepler}}
\newcommand{\vespa}{\texttt{vespa}}
\newcommand{\isochrones}{\texttt{isochrones}}
\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}

@@

But for \kepler, following this model has been largely impossible
because of the quantity and character of the flood of planet
candidates (thousands of mostly small-planet candidates around
relatively faint stars).  There have been a small number of
\kepler\ planets with masses measured by RVs
\citep[][e.g.]{Marcy:2014}, and significantly more that have been
confirmed as planets by measurement of transit timing variations
(TTVs) in multi-planet systems \citep[][e.g.]{}, but this still leaves
the vast majority of candidates unaccessible to dynamical
confirmation.

This situation has inspired the development of \emph{probabilistic
  validation} as a new approach to evaluating transit candidates.  The
principle of probabilistic validation is to demonstrate that all
conceivable astrophysical false positive scenarios are negligibly
likely to be the cause of a transit candidate signal compared to the
explanation of a planet transiting the presumed target star.  The
\verb|BLENDER| method pioneered this approach and has validated many
\kepler\ candidates, including .  More recently, the \verb|PASTIS| analysis
suite has been introduced \citep{Diaz:2014} and used to validate both
\kepler\ and \textit{CoRoT} candidates \citep{}.

While they have both proven useful for the purposes of validating
individual candidates of particular interest, neither \verb|BLENDER|
nor \verb|PASTIS| is designed for fully automated batch processing of
large numbers of candidates.  \citet{Morton:2012} describe a
computationally simpler planet validation procedure designed for
exactly such a purpose, based on the idea of describing eclipse light
curves as simple trapezoids and simulating realistic populations of
astrophysical false positives.  This procedure has also been used in
the literature to validate a number of \kepler\ planets, and has also
been applied to a number of candidates found by the \textit{K2}
mission \citep{Montet:2015,Becker:2015,Schlieder:2015}.  The code that
implements this procedure is now publicly available in the Python
module \vespa\footnote{\url{https://github.com/timothydmorton/vespa}}
\citep{vespa}.

@@

The goal of this work is to calculate the FPP for every KOI,
regardless of classification as CONFIRMED, CANDIDATE, or FALSE
POSITIVE.  As such, we begin with a list of \ntotal\ KOIs from the
cumulative table at the NASA Exoplanet Science Institute (NExScI)
Exoplanet Archive.  We then gather ancillary data and constraints from
various sources in order to enable the \vespa\ calculation:

@@

We were not able to successfully run \vespa\ on \nfail\ of the
\ntotal\ total KOIs. The various reasons for these failures are
detailed below.  
\begin{enumerate}
\item \nbadphot\ KOIs did not receive MCMC modeling.  Most of these
  (\nbadphotFP) are already designated FALSE POSITIVE at the archive,
  though \nbadphotCAND\ have CANDIDATE disposition.
\item \nbadrowemcmc\ KOIs did receive MCMC modeling but had unphysical fit
  results; e.g., negative $R_p/R_\star$ or best-fit impact parameter
  greater than $(1 + R_p/R_\star)$.
\item The host stars of \nbadstellar\ KOIs are not included in the
  \citetalias{Huber:2014} stellar property catalog, and thus were not
  included in this analysis.
\item \nbadsec\ KOIs have no \verb|koi_depth_err1| value on the archive,
  and thus had no weak secondary constraint and were left out of the
  calculations.
\item For \nbadtrapfit\ KOIs, the trapezoid MCMC fit did not converge.  The
  convergence criterion was for the autocorrelation time of the chain
  for each parameter to be shorter than 10\% of the total chain
  length.
\item For \nbadroche\ KOIs, the period of the planet was such that its
  orbit was implied to be within its host star's Roche limit.  This
  usually happens when the host star is estimated to be a giant, and
  these situations are nearly always false positives. 
\item Finally, \nbadother\ KOIs failed for as-yet undetermined
  reasons.  \nbadotherFP\ of these are already designated FALSE
  POSITIVE, \nbadotherCAND\ are CANDIDATE, and \nbadotherCONF\ is
  dispositioned as a confirmed planet.
\end{enumerate}
The numbers in this list correspond to the numbers in
the ``failure'' column in \tab{fpp}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Comparison with follow-up observations}
\sectlabel{comparisons} One of the difficulties with probabilistic
validation is that, by necessity, it is typically invoked when no
other method of follow-up confirmation is possible.  It can therefore
be difficult to find ways to compare the results of a calculation such
as \vespa\ to any known observational ground-truth, or to ``validate
the validations.''  However, because so much follow-up observational
effort has gone into \kepler\ candidates over the last few years,
there actually are two different datasets that do provide such
information for relatively small subsets of candidates.  Here, we
discuss the results of \vespa\ calculations for these candidates and
the impliciations for the reliability of the \vespa\ framework.

The first of these datasets is from \citet{Desert:2015}, who observe
the transits of 50 \kepler\ candidates with the Spitzer Space
Telescope in order to observationally constrain the FP rate.  The idea
behind these observations is that a blended EB false positive will
often show a color-dependent transit depth, and so comparing the
candidate depths measured by \kepler\ to those measured in the
infrared by Spitzer would give an idea of how likely a signal is to be
a false positive. The results of their analysis suggest that fewer
that about 8\% of their observed candidates are likely to be false
positives.  Of these 50 candidates, \vespa\ on its own calculates FPP
$\gt$ 0.1 for four.  Two of these (KOI-103.01 and KOI-248.02) are
systems with known significant TTVs; the other two (KOI-247.01 and
KOI-555.02) are likely false positives, with FPPs of 0.90 and 0.81,
respectively.  All but two of the remaining canidates have FPP $\lt$
0.01.

@@

In this work, we have calculated the astrophysical false positive
probability (FPP) for every \kepler\ object of interest (KOI), using
the publicly available Python module \vespa.  While the assumptions
behind this calculation are not necessarily valid for every KOI (see
\sect{results:fpp}), we have identified \nvalnew\ KOIs that have
reliable FPPs of $< 0.01$, resulting in validation of their planetary
nature at the 99\% confidence level, more than doubling the number of
confirmed \kepler\ exoplanets.  Among this set of reliable \vespa\
calculations are also \nfpnew\ new likely false positive KOIs,
although we note that there is a chance that some of these large FPPs
are due to unidentified or miscorrected transit timing variations.

The reliability of these calculations is dependent largely on the
results of \citet{Bryson:KSCI}, which quantify the probability that
the eclipse/transit signal is spatially coincident on the sky with the
presumed target star.  Without confirmation that the transit signal is
not coming from a significantly displaced source, the sky area used as
part of the prior for the false positive scenarios would need to be
significantly larger than the positional uncertainty value assumed in
this work (described in \sect{methods:fpp}).  Additionally, the
blended false positive scenarios \vespa\ considers are assumed to be
fully contained within the photometric aperture; this assumption would
also be broken if the source of the eclipse were significantly
displaced.

@@

The \kepler\ mission has demonstrated that space-based transiting
planet surveys identify planet candidates at a rate much faster than
traditional follow-up techniques can confirm them.  As a result, false
positive probability quantification techniques are now an integral
part of the landscape of exoplanet science.  While the present work is
the first large-scale demonstration of a fully automated validation
procedure, there is much progress still to be made.  For example,
there is currently no support within \vespa\ to calculate the FPP for
a candidate which has a \emph{specifically identified} but hitherto
unknown close companion.  Future development plans for the
\vespa\ package include support for this scenario, as well as other
improvements.  One of the most important of these will be to allow for
a contaminating EB to not be fully contained within the target
photometric aperture; that is, modeling the probability for
further-away stars to be EBs contributing only a small amount of their
flux to the target photometry, such as is the case for many of the
false positives identified via ephemeris matching by
\citet{Coughlin:2014}.  Full inclusion of this effect will allow for
even low-SNR candidates to receive confident \vespa\ analysis, which
is now limited only to KOIs for which confident pixel-level positional
analysis is possible.

>>>>

\newcommand{\nbadphot}{47} %EmptyPhotometry or MissingKOI
\newcommand{\nbadphotFP}{34}
\newcommand{\nbadphotCAND}{13}
\newcommand{\nbadrowemcmc}{233}
\newcommand{\nbadstellar}{74} %MissingStellar
\newcommand{\nbadsec}{8} %NoWeakSecondary
\newcommand{\nattempted}{8141}
\newcommand{\ntryfail}{112}
\newcommand{\nbadtrapfit}{36}
\newcommand{\nbadroche}{37}
\newcommand{\nbadother}{2}
\newcommand{\nbadotherFP}{1}
\newcommand{\nbadotherCAND}{1}
\newcommand{\nbadotherCONF}{1}

@@

\newcommand{\kepler}{\textit{Kepler}}
\newcommand{\vespa}{\texttt{vespa}}
\newcommand{\pastis}{\verb|PASTIS|}
\newcommand{\blender}{\verb|BLENDER|}
\newcommand{\isochrones}{\texttt{isochrones}}
\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}

@@

However, following this model has been largely impossible for \kepler\
because of the quantity and character of the planet candidates
(thousands of mostly small-planet candidates around relatively faint
stars).  There have been a small number of \kepler\ planets with
masses measured by RVs \citep[e.g.,][]{Marcy:2014}, and significantly
more that have been confirmed as planets by measurement of transit
timing variations (TTVs) in multi-planet systems
\citep[e.g.,][]{Ford:2012,Steffen:2012,Fabrycky:2012,Steffen:2013},
but this still leaves the vast majority of candidates unaccessible to
dynamical confirmation.

This situation has inspired the development of \emph{probabilistic
validation} as a new approach to evaluating transit candidates.  The
principle of probabilistic validation is to demonstrate that all
conceivable astrophysical false positive scenarios are negligibly
likely to be the cause of a transit candidate signal compared to the
explanation of a planet transiting the presumed target star.  The
\blender\ method pioneered this approach and has validated many
\kepler\ candidates
\citep[e.g.,][]{Borucki:2012,Kipping:2014,Torres:2015}.   More
recently, the \pastis\ analysis suite has been introduced
\citep{Diaz:2014} and used to validate both \kepler\ and
\textit{CoRoT} candidates \citep[e.g.,][]{Santerne:2014,Moutou:2014}.

While they have both proven useful for the purposes of validating
individual candidates of particular interest, neither \blender\ nor
\pastis\ is designed for fully automated batch processing of large
numbers of candidates.  An alternative validation approach for
candidates in multiple-planet systems has also been applied to a large
number of \kepler\ systems based on the general argument that it is
unlikely to see multiple false-positive signals in the same \kepler\
light curve \citep{Lissauer:2012}, resulting resulting in validations
of over 800 planets with 99\% confidence
\citep{Lissauer:2014,Rowe:2014}. However, this methodology differs
from the \blender/\pastis\ approach in two important ways: (a) it is
applicable only to planets in multi-planet systems, and (b) it relies on  
broad-brush general arguments rather than analyzing candidate signals
individually.

\citet{Morton:2012} describe a planet validation procedure that does
analyze individual signals, but is computationally simpler than
\blender\ or \pastis, designed explicitly for batch processing.  Based
on the idea of describing eclipse light curves as simple trapezoids
and simulating realistic populations of astrophysical false positives,
This procedure has also been used in the literature to validate a
number of \kepler\ planets
\citep[e.g.,][]{Muirhead:2012,Dawson:2012,Swift:2013}, and has also
been applied to a number of candidates found by the \textit{K2}
mission \citep{Montet:2015,Becker:2015}.  The code that implements
this procedure is publicly available in the Python module
\vespa\footnote{\url{https://github.com/timothydmorton/vespa}}
\citep{vespa}.

@@

The goal of this work is to calculate the FPP for every KOI,
regardless of classification as CONFIRMED, CANDIDATE, or FALSE
POSITIVE.  As such, we begin with a list of \ntotal\ KOIs from the
Q1-Q17 (DR24) table at the NASA Exoplanet Science Institute (NExScI)
Exoplanet Archive (the most recent available uniform catalog).  We
then gather ancillary data and constraints from various sources in
order to enable the \vespa\ calculation:

@@

We were not able to successfully run \vespa\ on all of the KOIs. The 
various reasons for these failures are detailed below.  
\begin{enumerate}
\item Some KOIs did not receive MCMC modeling.  Most of these
  have been are already designated FALSE POSITIVE at the archive.
\item A number of KOIs (\nbadrowemcmc) did receive MCMC modeling but had unphysical fit
  results; e.g., negative $R_p/R_\star$ or best-fit impact parameter
  greater than $(1 + R_p/R_\star)$.  These KOIs were left out of the \vespa calculations.
\item The host stars of \nbadstellar\ KOIs are not included in the
  \citetalias{Huber:2014} stellar property catalog, and thus were not
  included in this analysis.
\item \nbadsec\ KOIs have no \verb|koi_depth_err1| value on the archive,
  and thus had no weak secondary constraint and were left out of the
  calculations.
\item For \nbadtrapfit\ KOIs, the trapezoid MCMC fit did not converge.  The
  convergence criterion was for the autocorrelation time of the chain
  for each parameter to be shorter than 10\% of the total chain
  length.
\item For \nbadroche\ KOIs, the period of the planet was such that its
  orbit was implied to be within its host star's Roche limit.  This
  usually happens when the host star is estimated to be a giant, and
  these situations are nearly always false positives. 
\item Finally, \nbadother\ KOIs failed for as-yet undetermined
  reasons, one of which is a FALSE POSITIVE (371.02) and one is a 
  candidate (2433.06).
\end{enumerate}
The numbers in this list correspond to the numbers in
the ``failure'' column in \tab{fpp}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Comparison with follow-up surveys}
\sectlabel{comparisons}

One of the difficulties with probabilistic
validation is that, by necessity, it is typically invoked when no
other method of follow-up confirmation is possible.  It can therefore
be difficult to find ways to compare the results of a calculation such
as \vespa\ to any known observational ground-truth, or to ``validate
the validations.''  However, because so much follow-up observational
effort has gone into \kepler\ candidates over the last few years,
there actually are two different datasets that do provide such
information for relatively small subsets of candidates.  Here, we
discuss the results of \vespa\ calculations for these candidates and
the impliciations for the reliability of the \vespa\ framework.

The first of these datasets is from \citet{Desert:2015}, who observe
the transits of 50 \kepler\ candidates with the Spitzer Space
Telescope in order to observationally constrain the FP rate.  The idea
behind these observations is that a blended EB false positive will
often show a color-dependent transit depth, and so comparing the
candidate depths measured by \kepler\ to those measured in the
infrared by Spitzer would give an idea of how likely a signal is to be
a false positive. The results of their analysis suggest that fewer
than 8\% of their observed candidates are likely to be false
positives.  Of these 50 candidates, \vespa\ on its own calculates FPP
$>$ 0.1 for four.  Two of these (KOI-103.01 and KOI-248.02) are
systems with known significant TTVs; the other two (KOI-247.01 and
KOI-555.02) are likely false positives, with FPPs of 0.90 and 0.81,
respectively.  For all but two of the remaining candidates, \vespa\
gives FPP $< 0.01$.  The false positive rates calculated in
\citet{Desert:2015} are also based on probabilistic arguments very
similar to \citet{Morton:2011b} and thus are not quite candidates for
``ground-truth'' comparison, but the lack of transit chromaticity and
the agreements between these two independent studies are certainly
supportive of the \vespa\ results.

A much more independent and powerful test data set has recently become
available in the work of \citet{Santerne:2015}, which presented the
results of a long-term RV-monitoring campaign targeting 129 \kepler\
giant-planet candidates.  Of these, they confirm 45 to be planets and
identify 48 as eclipsing binaries, 15 as contaminating EBs (CEBs), and
3 as brown dwarfs.  They are unable to determine the nature of the
remaining 18.  These results imply a relatively high FP rate among
giant-planet candidates, possibly near 50\%, which sounds potentially
concerning.  However, when we look at our results on this sample, we
see very good agreement between the \vespa\ predictions and the RV-
detected ``ground truth'' (\Tab{santerne}).  Notably, confirmed
planets show a mean FPP of about 9\% and a median of only 0.02\%,
while confirmed EBs (CEBs) show a mean FPP of 78\% (83\%) and a median
of 95\% (99.5\%).  The three brown dwarfs also show low FPPs, which is
understandable because \vespa\ doesn't pretend to predict anything
about the mass of the companion, and BDs are essentially the same size
as giant planets.  It is also instructive to investigate the four
cases where \vespa\ computes high FPPs ($>$0.3) for confirmed planets.
Three of these (KOI-377.01, KOI-1426.02 and KOI-1474.01) have
significant TTVs, and one is a grazing eclipse (KOI-614.01).
KOI-1474.01 is also on a highly eccentric orbit \citep{Dawson:2012},
in addition to its TTVs, which also contributes to its high FPP.
Overall, this comparison powerfully demonstrates the reliability of
the \vespa\ calculation, showing that even in a population of
candidates that include many false positives, it is able to
effectively identify which are true planets and which are not.

\begin{deluxetable}{ccc}
\tabletypesize{\scriptsize}
\tablecaption{\vespa--calculated FPPs of the \\Santerne (2015) RV sample
  \tablabel{santerne}}
\tablehead{\colhead{nature} & \colhead{mean FPP} & \colhead{median FPP}}
\startdata
planet & 0.0892 & 0.0002 \\
BD & 0.0035 & 0.0038 \\
EB & 0.7764 & 0.9450 \\
CEB & 0.8326 & 0.9950 \\
? & 0.2393 & 0.0032
\enddata
\end{deluxetable}

@@

In this work, we have calculated the astrophysical false positive
probability (FPP) for every \kepler\ object of interest (KOI) in the
Q1-Q17 (DR24) table, using the publicly available Python module
\vespa, which implements the procedure introduced in
\citet{Morton:2012}, with improvements in stellar parameter modeling
and the inclusion of new ``double-period'' false positive scenarios.
While the assumptions behind this calculation are not necessarily
valid for every KOI (see \sect{results:fpp}), we have identified
\nvalnew\ KOIs that have reliable FPPs of $< 0.01$, resulting in
validation of their planetary nature at the 99\% confidence level,
more than doubling the number of confirmed \kepler\ exoplanets.  Among
this set of reliable \vespa\ calculations are also \nfpnew\ new likely
false positive KOIs, although we note that some of these may be due to
unidentified or miscorrected transit timing variations.

These calculations depend largely on the results of
\citet{Bryson:KSCI}, which quantify the probability that the
eclipse/transit signal is spatially coincident on the sky with the
presumed target star.  Without confirmation that the transit signal is
not coming from a significantly displaced source, the sky area used as
part of the prior for the false positive scenarios would need to be
significantly larger than the positional uncertainty value assumed in
this work (described in \sect{methods:fpp}).  Additionally, the
blended false positive scenarios \vespa\ considers are assumed to be
fully contained within the photometric aperture; this assumption would
also be broken if the source of the eclipse were significantly
displaced.

@@

The \kepler\ mission has demonstrated that space-based transiting
planet surveys identify planet candidates at a rate much faster than
traditional follow-up techniques can confirm them.  As a result, false
positive probability quantification techniques are now an integral
part of the landscape of exoplanet science.  While the present work is
the first large-scale demonstration of a fully automated validation
procedure, there is much progress still to be made.  For example,
there is currently no support within \vespa\ to calculate the FPP for
a candidate which has a \emph{specifically identified} but previously
unknown close companion.  Future development plans for the
\vespa\ package include support for this scenario, as well as other
improvements.  One of the most important of these will be to allow for
a contaminating EB to not be fully contained within the target
photometric aperture; that is, modeling the probability for
further-away stars to be EBs contributing only a small amount of their
flux to the target photometry, such as is the case for many of the
false positives identified via ephemeris matching by
\citet{Coughlin:2014}.  Full inclusion of this effect will allow for
even low-SNR candidates to receive confident \vespa\ analysis, which
is now limited only to KOIs for which confident pixel-level positional
analysis is possible.
